---
title: "4-TTH"
author: "bernard-liew"
date: "2021-06-23"
output: 
  workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Load package

```{r, message=F, warning=F}

rm (list = ls())
# Helper
library (tidyverse)
library (skimr)
library (cowplot)
library (ggpubr)
library (flextable)
library (officer)
library (arsenal)
library (rstatix)
library (kableExtra)
library (gtsummary)
library (ggdendro)
# BN
library (bnlearn)
library (Rgraphviz)

#SEM
library (lavaan)
library (semPlot)
library (semptools)
library (parameters)
library (SEMsens)

# Cluster 
library (cluster)
library (factoextra)
library (fastcluster)
library (clValid)
```

```{r include = FALSE}

res <- readRDS("output/res.RDS")
list2env(res,globalenv())
```


# Load data

This is an imputed dataset

```{r}

df_base <- readRDS("output/df_base.RDS")

#df_base_imp <-mice (df_base$raw, m = 5, maxit = 30, method = "rf")

df <- mice::complete (df_base$imp) %>%
  select (-c(strn, flexible, freedom)) %>%
  mutate (sex = factor (sex, labels = c("0", "1"))) %>%
  mutate_at (vars(fitness:balance), as.numeric)

df <-df %>%
  mutate (duration = fct_collapse(duration,
                                  "0" = c("1","2", "3"),
                                  "1" = "4",
                                  "2" = "5"))
```

## Preview data

```{r}
tbl_summary(df) 
```

```{r}
df %>%
  kable () %>%
    kable_styling() %>%
    scroll_box(width = "700px", height = "500px")
```



# Cluster analysis

Performs cluster analysis using Gower's distance and `complete` linkage.

```{r, eval = FALSE}
dd <- daisy(df, metric="gower")

res.hc <- fastcluster::hclust(d = dd, method = "complete")
```


## Validate

```{r, eval = FALSE}
plot (res.hc)

dd_mat <- as.matrix(dd)
intern <- clValid(dd_mat, 2:6, clMethods=c("hierarchical"), method = "complete",
                   validation="internal")
```

## Plot

```{r, eval = FALSE}
dend <- res.hc %>%
  as.dendrogram %>%
  set("branches_k_color", k = 2) %>% set("branches_lwd", 0.7) %>%
  set("labels_cex", 12) %>% set("labels_colors", k = 2) %>%
  set("leaves_pch", 19) %>% set("leaves_cex", 0.5) 
ggd1 <- as.ggdend(dend)
f <- ggplot(ggd1, horiz = FALSE, labels = FALSE) + 
  ylab ("Gower's distance")


tiff ("manuscript/fig2.tiff", height = 8, width = 12, units = "in", res = 100)
ggdendro::ggdendrogram(res.hc, 
                       labels = FALSE, 
                       leaf_labels = FALSE) + 
  theme_cowplot() + 
  xlab ("Observation number") + 
  ylab ("Gower's distance")
dev.off ()
```

## Select optimal clusters

Selected two from visual inspection, and collapsed two variables `flex_mob` and `duration` into binary variables.

```{r}

grp <- cutree(res.hc, k = 2)


df_grp <- df %>%
  mutate (grp = factor (grp)) 

tab1 <- tableby(grp ~ ., data=df_grp)
summary (tab1, text=TRUE)
```


# Descriptives

```{r}
meanNsd_transform <- function (x) {
  m <- round (as.numeric (x[[1]][1], 2))
  s <- round (as.numeric (x[[1]][2], 2))
  m_s <- paste0(m, "(", s, ")")
  return (m_s)
}

p_round <- function (x) {
  m <- round (as.numeric (x[[1]][1], 3))
  if (m < 0.001) m = "<0.001"
  return (m)
}
tab1 <- tableby (grp ~ ., data=df_grp, digits = 2, digits.p = 2) %>%
  as.data.frame() %>%
  filter (!term %in% c("Nmiss", "range")) %>%
  select (-c(group.term:term, variable.type)) 
tab2 <-  tab1[-c (6, 8, 13, 15, 17, 23, 25, 27, 29, 31, 33, 35),]

tab2 <- tab2%>% mutate_all(na_if,"")
tab2$label <- tab1$label[!grepl ("Mean", tab1$label)]
tab2[,tab2[,2]==""] <- NA
tab2[,tab2[,3]==""] <- NA
tab2[,tab2[,5]==""] <- NA
for(row in 1:nrow(tab2)) {
    tab2[row, 2] <- meanNsd_transform (tab2[row, 2])
    tab2[row, 3] <- meanNsd_transform (tab2[row, 3])
    tab2[row, 4] <- meanNsd_transform (tab2[row, 4])
    tab2[row, 6] <- p_round (tab2[row, 6])
  }
colnames (tab2) <- c ("Variables", 
                      "Group 1",
                      "Group 2",
                      "Total",
                      "Statistic",
                      "P value")

tab2$`Group 1` <- unlist(tab2$`Group 1`)
tab2$`Group 2` <- unlist(tab2$`Group 2`)
tab2$`Total` <- unlist(tab2$`Total`)
# my_path <- paste0("manuscript/table_1",
#                   ".docx")
# 
# ft <- flextable(tab2) %>%
#   set_caption(caption = " Table 1.Baseline descriptive characteristics of cohort") %>%
#   autofit()
# 
# my_doc <- read_docx()  %>%
#   body_add_flextable(ft)
# 
# print (my_doc, target = my_path)
```

```{r}
tab2 %>%
  kbl() %>%
  kable_styling()
```

# Preprocess

Scale numeric variables to mean -0 and SD - 1

```{r}
df_sem <- df_grp %>%
  #select (-c(strn, flexible, freedom)) %>%
  mutate_if(is.numeric, scale, center = TRUE, scale = TRUE) %>%
  mutate_if(is.factor, as.numeric)%>%
  mutate (sex = sex -1)

```


# CFA

Test latent model

```{r}
cfa_form0 <- "
  # latent
  physical =~ abds_ms + ext_ms + flex_mob 
  pain =~ lbp + legp + duration
  psych =~ ipq + fabq + ases
  funx =~ fitness + endure + balance 
  
  duration~~legp
  ipq~~flex_mob
  lbp~~flex_mob
"


cfa0 <- lavaan::cfa (cfa_form0, 
           data = df_sem,
           ordered = c("flex_mob", "duration"),
           std.lv = FALSE, 
           auto.fix.first = TRUE, 
           estimator = "WLSMV",
           se = "robust.sem")

res1 <- residuals(cfa0, type = "cor")$cov %>% 
  cor_gather() %>%
  mutate (form = paste0(var1, "~~", var2))  %>%
  filter (abs(cor) > 0.10) %>%
  arrange (abs(cor))

fitMeasures(cfa0, fit.measures = c("rmsea", "pvalue", "cfi", "tli", "srmr", "nnfi"))
```

## Plot

```{r}
 semPaths (cfa0, 
          what = "paths",
          whatLabels = "est",
          nCharNodes = 0,
          layout = "tree",
          label.cex = 1.25,
          sizeMan = 6,
          thresholds = FALSE,
          intercepts = FALSE)




```

```{r, eval = FALSE}
my_label_list <- list(list(node = "funx", to = "activity"))
f <- semPaths (cfa0, 
          what = "paths",
          whatLabels = "est",
          nCharNodes = 0,
          layout = "tree",
          label.cex = 1,
          sizeMan = 6,
          residuals = FALSE, 
          thresholds = FALSE,
          intercepts = FALSE)
f2 <- change_node_label(f, my_label_list)
tiff ("manuscript/fig3.tiff", height = 8, width = 12, units = "in", res = 100)
plot (f2)
dev.off ()
```


# SEM analysis

Form three datasets.

`df_sem1` = cluster 1
`df_sem2` = cluster 2
`df_sem`  = whole group data

```{r, eval = FALSE}
df_sem1 <- df_grp %>%
  filter (grp ==1) %>%
  select (-grp) %>%
  mutate_if(is.numeric, scale, center = TRUE, scale = TRUE) %>%
  mutate_if(is.factor, as.numeric) %>%
  mutate (sex = sex - 1)
  

df_sem2 <- df_grp %>%
  filter (grp ==2) %>%
  select (-grp) %>%
  mutate_if(is.numeric, scale, center = TRUE, scale = TRUE) %>%
  mutate_if(is.factor, as.numeric)%>%
  mutate (sex = sex - 1)


```


## Bayesian networks

Creates 3 datasets for bayesian networks. Need to initialize latent variables and treat them as missing data for structural learning.

```{r, eval = FALSE}

num_vars <- names(df_sem)[! names(df_sem) %in% c("flex_mob", "grp", "duration", "sex")]

df.bn <- as.data.frame (df_sem) %>%
  select (-grp) %>%
  mutate (pain = as.numeric (NA),
          psych = as.numeric(NA),
          funx = as.numeric (NA),
          physical = as.numeric(NA))

df.bn1 <- as.data.frame (df_sem1) %>%
  mutate (pain = as.numeric (NA),
          psych = as.numeric(NA),
          funx = as.numeric (NA),
          physical = as.numeric(NA))%>%
  mutate_if (is.factor, as.numeric)

df.bn2 <- as.data.frame (df_sem2) %>%
  mutate (pain = as.numeric (NA),
          psych = as.numeric(NA),
          funx = as.numeric (NA),
          physical = as.numeric(NA))%>%
  mutate_if (is.factor, as.numeric)


```


### Create blacklist and whitelist

Whitelist are from the latent to manifest variables.

Blacklist are that manifest variables can only point from their respective latent variables.

```{r, eval = FALSE}

demo.var = grep("age|sex", colnames (df.bn), value = TRUE)
pain.var = grep("legp|lbp|duration", colnames (df.bn), value = TRUE)
psych.var = grep("ases|fabq|ipq", colnames (df.bn), value = TRUE)
phys.var = grep("ext|abds|flex_", colnames (df.bn), value = TRUE)
fx.var = grep("fitness|endure|balance", colnames (df.bn), value = TRUE)
outcome.var = grep("odi", colnames (df.bn), value = TRUE)
others.var = setdiff (names(df.bn), c(demo.var, pain.var, psych.var, phys.var, fx.var, outcome.var))

pair_var <- expand.grid(from = names (df.bn),
                        to = names (df.bn)) %>%
  rownames_to_column()

tiers_keep <- pair_var %>%
  filter (!(grepl (paste0(c(pain.var, psych.var, phys.var, fx.var, demo.var), collapse = "|"),to))) %>%
  filter (!(grepl (paste0(c(pain.var, psych.var, phys.var, fx.var, outcome.var), collapse = "|"),from))) 

  
bl <- anti_join(pair_var, tiers_keep, by = "rowname")  %>%
  filter (from != to) %>%
  select (-rowname) %>%
  mutate_all(as.character) %>% 
  filter (from !="grp") %>% 
  filter (to !="grp") 

latent1 <- expand.grid(from = "pain",
                        to = pain.var)
latent2 <- expand.grid(from = "psych",
                        to = psych.var)
latent3 <- expand.grid(from = "physical",
                        to = phys.var)
latent4 <- expand.grid(from = "funx",
                        to = fx.var)
wl <- latent1 %>%
  bind_rows(latent2) %>%
  bind_rows(latent3)%>%
  bind_rows(latent4)%>%
  mutate_all (as.character)


```

```{r, eval = FALSE}
writexl::write_xlsx(list (blacklist = bl, 
                          whitelist = wl), 
                    "manuscript/sm_black_white_list.xlsx")
```


### Structural learning

```{r, eval= FALSE}

####################### Full dataset ####################### ##################

start = bnlearn::bn.fit(bnlearn::empty.graph(names(df.bn)), df.bn)
start$pain = list(coef = 0, sd = 1)
start$physical = list(coef = 0, sd = 1)
start$psych = list(coef = 0, sd = 1)
start$funx = list(coef = 0, sd = 1)

m0 <- bnlearn::structural.em(df.bn, 
                       start = start,
                       maximize = "hc",
                       maximize.args = list(blacklist = bl,
                                            whitelist = wl))


####################### subgroup 1 ####################### ##################
start = bnlearn::bn.fit(bnlearn::empty.graph(names(df.bn1)), df.bn1)
start$pain = list(coef = 0, sd = 1)
start$physical = list(coef = 0, sd = 1)
start$psych = list(coef = 0, sd = 1)
start$funx = list(coef = 0, sd = 1)

m1 <- bnlearn::structural.em(df.bn1, 
                       start = start,
                       maximize = "hc",
                       maximize.args = list(blacklist = bl,
                                            whitelist = wl))



####################### subgroup 2 ####################### ##################
start = bnlearn::bn.fit(bnlearn::empty.graph(names(df.bn2)), df.bn2)
start$pain = list(coef = 0, sd = 1)
start$physical = list(coef = 0, sd = 1)
start$psych = list(coef = 0, sd = 1)
start$funx = list(coef = 0, sd = 1)

m2 <- bnlearn::structural.em(df.bn2, 
                       start = start,
                       maximize = "hc",
                       maximize.args = list(blacklist = bl,
                                            whitelist = wl))


## Plot

par (mfrow = c(2,2), mar = c(1, 1, 1, 1))
plot (m0)
plot (m1)
plot (m2)

```

## BN to SEM 

```{r}
latent_var <- c("age", "odi", "psych", "physical", "funx", "pain", "sex")

edge_2_form <- function (edge, vars2form, otherform) {
  
  edge <- as.data.frame(edge)
  
  edge <- edge[edge$to %in%vars2form, ] 
  form_vec <- edge %>%
    unite ("form", to, from, sep =  "~") %>%
    pull()
  
  paste0 (form_vec,  collapse = "\n")
  form  = paste0(otherform, paste0 (form_vec,  collapse = "\n"),  collapse = "\n")
  return (form)
}


```


### Whole cohort

```{r}
edge = arcs (m0)

sem_form0 <- edge_2_form(edge = edge,
            vars2form = latent_var,
            otherform = cfa_form0)

m0_sem <- lavaan::sem (sem_form0,  
                       data = df_sem,
                       ordered = c("flex_mob", "duration"),
                       std.lv = FALSE, 
                       auto.fix.first = TRUE, 
                       estimator = "WLSMV",
                       se = "robust.sem")
fitMeasures(m0_sem, fit.measures = c("rmsea", "pvalue", "cfi", "tli", "srmr", "nnfi"))
```

```{r, echo  = FALSE}
p1 <- semPaths (m0_sem, 
          what = "path", 
          layout = "tree3", 
          curvePivot = TRUE, 
          residuals = FALSE,
          intercept = FALSE,
          nCharNodes = 0,
          thresholds = FALSE,
          exoCov = FALSE,
          #fixedStyle = c(adjustcolor( "white", alpha.f = 0), 0),
          rotation = 2,
          label.cex = 1,
          curve = TRUE,
          freeStyle = c("black", 1))
```

```{r}
my_label_list <- list(list(node = "funx", to = "activity"),
                      list (node = "sex", to = "gender"))

indicator_order <- c("flex_mob", "abds_ms", "ext_ms",
                     "lbp", "legp", "duration",
                     "ipq", "fabq", "ases",
                     "fitness", "endure", "balance",
                     "age", "sex", "odi")

indicator_factor <- c("physical", "physical", "physical",
                     "pain", "pain", "pain",
                     "psych", "psych", "psych",
                     "funx", "funx", "funx",
                     "age", "sex", "odi")

factor_layout <- matrix(c(NA,  "age", NA,    NA,  NA,
                          "funx", NA, NA, "pain", NA,
                          NA, NA, NA, NA, NA, 
                          NA, "sex", NA, NA, "odi",
                          NA, NA, NA, NA, NA, 
                          "physical", NA, NA,  "psych", NA), byrow = TRUE, 6, 5)

factor_point_to <- matrix(c(NA,   "left",NA,    NA,  NA,
                          "left", NA, NA, "up", NA,
                          NA, NA, NA, NA, NA, 
                          NA, "left", NA, NA, "left",
                          NA, NA, NA, NA, NA, 
                          "left", NA, NA,  "down", NA), byrow = TRUE, 6, 5)


indicator_push <- c(funx = 2,
                      physical = 2,
                      pain = 3,
                      psych = 3,
                    age = 0,
                    sex = 0,
                    odi = 0)

indicator_spread <- c(funx = 4,
                      physical = 4,
                      pain = 3,
                      psych = 3,
                      age = 0,
                      sex = 0,
                      odi = 0)
p2 <- set_sem_layout(p1, 
                     indicator_order = indicator_order,
                     indicator_factor = indicator_factor,
                     factor_layout = factor_layout,
                     factor_point_to = factor_point_to,
                     indicator_push = indicator_push,
                     indicator_spread = indicator_spread) %>%
  mark_sig(object = m0_sem) %>%
  change_node_label(my_label_list)
plot(p2)


```


```{r, eval = FALSE}
tiff ("manuscript/fig3.tiff", height = 12, width = 20, units = "in", res = 100)
plot(p2)
dev.off ()
```

```{r, eval = FALSE}

param_ex <- model_parameters(m0_sem, standardize = TRUE,
                 component = c("regression", "loading")) %>%
  rename (DV = To,
          IV = From,
          Zval = z,
          Pval = p) %>%
  mutate (Component = ifelse (Component == "Loading", "Latent", Component)) %>%
  mutate (Sig = ifelse (Pval < 0.05, "s", "ns")) %>%
  select (- c(Operator)) %>%
  mutate_if(is.numeric, round, 3)

# Export lavaan table
my_path <- paste0("manuscript/sm_table1", 
                  ".docx")
ft <- flextable(param_ex) %>%
  set_caption(caption = " Table 1.Parameter estimates for whole cohort") %>%
  autofit() 
my_doc <- read_docx()  %>% 
  body_add_flextable(ft)
print (my_doc, target = my_path)
```

### Subgroup 1

```{r, eval = FALSE}
edge = arcs (m1)

sem_form1 <- edge_2_form(edge = edge,
            vars2form = latent_var,
            otherform = cfa_form0)

m1_sem <- lavaan::sem (sem_form1,
                       data = df_sem1,
                       ordered = c("flex_mob", "duration"),
                       std.lv = FALSE, 
                       auto.fix.first = TRUE, 
                       estimator = "WLSMV",
                       se = "robust.sem")
summary (m1_sem)

fitMeasures(m1_sem, fit.measures = c("rmsea", "pvalue", "cfi", "tli", "srmr", "nnfi"))
```



```{r, echo  = FALSE}
p1 <- semPaths (m1_sem, 
          what = "path", 
          layout = "tree3", 
          curvePivot = FALSE, 
          residuals = FALSE,
          intercept = FALSE,
          nCharNodes = 0,
          thresholds = FALSE,
          exoCov = FALSE,
          #fixedStyle = c(adjustcolor( "white", alpha.f = 0), 0),
          rotation = 2,
          label.cex = 1,
          curve = TRUE,
          freeStyle = c("black", 1))
```

```{r}

p2 <- set_sem_layout(p1, 
                     indicator_order = indicator_order,
                     indicator_factor = indicator_factor,
                     factor_layout = factor_layout,
                     factor_point_to = factor_point_to,
                     indicator_push = indicator_push,
                     indicator_spread = indicator_spread) %>%
  mark_sig(object = m1_sem)%>%
  change_node_label(my_label_list)
plot(p2)
```

```{r, eval = FALSE}
tiff ("manuscript/fig4.tiff", height = 12, width = 20, units = "in", res = 100)
plot(p2)
dev.off ()
```

```{r, eval = FALSE}
param_ex <- model_parameters(m1_sem, standardize = TRUE,
                 component = c("regression", "loading")) %>%
  rename (DV = To,
          IV = From,
          Zval = z,
          Pval = p) %>%
  mutate (Component = ifelse (Component == "Loading", "Latent", Component)) %>%
  mutate (Sig = ifelse (Pval < 0.05, "s", "ns")) %>%
  select (- c(Operator)) %>%
  mutate_if(is.numeric, round, 3)

# Export lavaan table
my_path <- paste0("manuscript/sm_table2", 
                  ".docx")
ft <- flextable(param_ex) %>%
  set_caption(caption = " Table 2.Parameter estimates for subgroup 1") %>%
  autofit() 
my_doc <- read_docx()  %>% 
  body_add_flextable(ft)
print (my_doc, target = my_path)
```

### Subgroup 2

```{r, eval = FALSE}
edge = arcs (m2)

sem_form2 <- edge_2_form(edge = edge,
            vars2form = latent_var,
            otherform = cfa_form0)

m2_sem <- lavaan::sem (sem_form2,  data = df_sem2, ordered = c("flex_mob", "duration"))
summary (m2_sem )

fitMeasures(m2_sem , fit.measures = c("rmsea", "pvalue", "cfi", "tli", "srmr", "nnfi"))
```

```{r, echo  = FALSE}
p1 <- semPaths (m2_sem, 
          what = "path", 
          layout = "tree3", 
          curvePivot = FALSE, 
          residuals = FALSE,
          intercept = FALSE,
          nCharNodes = 0,
          thresholds = FALSE,
          exoCov = FALSE,
          #fixedStyle = c(adjustcolor( "white", alpha.f = 0), 0),
          rotation = 2,
          label.cex = 1,
          curve = TRUE,
          freeStyle = c("black", 1))
```

```{r, eval = FALSE}

p2 <- set_sem_layout(p1, 
                     indicator_order = indicator_order,
                     indicator_factor = indicator_factor,
                     factor_layout = factor_layout,
                     factor_point_to = factor_point_to,
                     indicator_push = indicator_push,
                     indicator_spread = indicator_spread) %>%
  mark_sig(object = m2_sem)%>%
  change_node_label(my_label_list)
plot(p2)
```

```{r, eval = FALSE}
tiff ("manuscript/fig5.tiff", height = 12, width = 20, units = "in", res = 100)
plot(p2)
dev.off ()
```

```{r, eval = FALSE}
param_ex <- model_parameters(m2_sem, standardize = TRUE,
                 component = c("regression", "loading")) %>%
  rename (DV = To,
          IV = From,
          Zval = z,
          Pval = p) %>%
  mutate (Component = ifelse (Component == "Loading", "Latent", Component)) %>%
  mutate (Sig = ifelse (Pval < 0.05, "s", "ns")) %>%
  select (- c(Operator)) %>%
  mutate_if(is.numeric, round, 3)
# Export lavaan table
my_path <- paste0("manuscript/sm_table3", 
                  ".docx")
ft <- flextable(param_ex) %>%
  set_caption(caption = " Table 3.Parameter estimates for subgroup 2") %>%
  autofit() 
my_doc <- read_docx()  %>% 
  body_add_flextable(ft)
print (my_doc, target = my_path)
```


# Save data

```{r, eval = FALSE}
res <- list (m0 = m0,
             m1 = m1,
             m2 = m2,
             m0_sem = m0_sem,
             m1_sem  = m1_sem,
             m2_sem  = m2_sem,
             df = df,
             df_grp = df_grp,
             df_sem = df_sem,
             df_sem1 = df_sem1,
             df_sem2 = df_sem2,
             df.bn = df.bn,
             df.bn1 = df.bn1,
             df.bn2 = df.bn2,
             dd = dd,
             res.hc = res.hc)

saveRDS (res,
         "output/res.RDS")
```

