---
title: "4-TTH"
author: "bernard-liew"
date: "2021-06-23"
output: 
  workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

# Load package

```{r, message=F, warning=F}

rm (list = ls())
# Helper
library (tidyverse)
library (skimr)
library (cowplot)
library (ggpubr)
library (flextable)
library (officer)
library (arsenal)
library (rstatix)
library (kableExtra)

# BN
library (bnlearn)
library (Rgraphviz)

#SEM
library (lavaan)
library (semPlot)
library (semptools)

# Cluster 
library (cluster)
library (factoextra)
library (fastcluster)
```

# Load data

This is an imputed dataset

```{r}

df <- mice::complete (readRDS("output/df_base.RDS")$imp) %>%
  select (-c(trajec, bmi, sbt)) %>%
  mutate (sex = factor (sex))%>%
  mutate_at (vars(fitness:freedom), as.numeric)
```

## Preview data

```{r}
skim (df)
```

```{r}
df %>%
  kable () %>%
    kable_styling() %>%
    scroll_box(width = "700px", height = "500px")
```



# Cluster analysis

Performs cluster analysis using Gower's distance and `ward.D2` linkage.

```{r}
dd <- daisy(df, metric="gower")

res.hc <- fastcluster::hclust(d = dd, method = "ward.D2")
plot (res.hc)
```

## Select optimal clusters

Selected two from visual inspection, and collapsed two variables `flex_mob` and `duration` into binary variables.

```{r}

grp <- cutree(res.hc, k = 2)


df_grp <- df %>%
  mutate (grp = factor (grp)) %>%
  mutate (flex_mob = fct_collapse(flex_mob,
                                  "0" = c("0","1", "2"),
                                  "1" = c("3", "4"))) %>%
  mutate (duration = fct_collapse(duration,
                                  "0" = c("1","2", "3"),
                                  "1" = c("4", "5")))

tab1 <- tableby(grp ~ ., data=df_grp)
summary (tab1, text=TRUE)
```


# Preprocess

Scale numeric variables to mean -0 and SD - 1

```{r}
df_sem <- df_grp %>%
  mutate_if(is.numeric, scale, center = TRUE, scale = TRUE) %>%
  mutate_if(is.factor, as.numeric)

```


# CFA

Test latent model

```{r}
cfa_form0 <- "
  # latent
  physical =~ flex_mob + abds_ms + ext_ms
  pain =~ lbp + legp + duration
  psych =~ ipq + fabq + ases
  funx =~ fitness + endure + balance 
  
"


m0 <- cfa (cfa_form0, 
           data = df_sem,
           ordered = c("flex_mob", "duration"))

fitMeasures(m0, fit.measures = c("rmsea", "pvalue", "cfi", "tli", "srmr", "nnfi"))
```

## Plot

```{r}
semPaths (m0, 
          what = "paths",
          whatLabels = "est",
          nCharNodes = 0,
          layout = "tree",
          label.cex = 1.5,
          sizeMan = 6,
          intercepts = FALSE)
```

# SEM analysis

Form three datasets.

`df_sem1` = cluster 1
`df_sem2` = cluster 2
`df_sem`  = whole group data

```{r}
df_sem1 <- df_grp %>%
  filter (grp ==1) %>%
  select (-grp) %>%
  mutate_if(is.numeric, scale, center = TRUE, scale = TRUE) %>%
  mutate_if(is.factor, as.numeric)
  

df_sem2 <- df_grp %>%
  filter (grp ==2) %>%
  select (-grp) %>%
  mutate_if(is.numeric, scale, center = TRUE, scale = TRUE) %>%
  mutate_if(is.factor, as.numeric)

```


## Bayesian networks

Creates 3 datasets for bayesian networks. Need to initialize latent variables and treat them as missing data for structural learning.

```{r}

df.bn <- as.data.frame (df_sem) %>%
  select (-grp) %>%
  mutate_if(is.numeric, scale, center = TRUE, scale = TRUE) %>%
  mutate (pain = as.numeric (NA),
          psych = as.numeric(NA),
          funx = as.numeric (NA),
          phys = as.numeric(NA))%>%
  mutate_if (is.factor, as.numeric)%>%
  select (-c(strn, flexible, freedom, sex, odi))

df.bn1 <- as.data.frame (df_sem1) %>%
  mutate_if(is.numeric, scale, center = TRUE, scale = TRUE) %>%
  mutate (pain = as.numeric (NA),
          psych = as.numeric(NA),
          funx = as.numeric (NA),
          phys = as.numeric(NA))%>%
  mutate_if (is.factor, as.numeric)%>%
  select (-c(strn, flexible, freedom, sex, odi))

df.bn2 <- as.data.frame (df_sem2) %>%
  mutate_if(is.numeric, scale, center = TRUE, scale = TRUE) %>%
  mutate (pain = as.numeric (NA),
          psych = as.numeric(NA),
          funx = as.numeric (NA),
          phys = as.numeric(NA))%>%
  mutate_if (is.factor, as.numeric)%>%
  select (-c(strn, flexible, freedom, sex, odi))

```


### Create blacklist and whitelist

Whitelist are from the latent to manifest variables
Blacklist are that manifest variables can only point from their respective latent variables.

```{r}

demo.var = grep("age|grp", colnames (df.bn1), value = TRUE)
pain.var = grep("duration|legp|lbp", colnames (df.bn1), value = TRUE)
psych.var = grep("ases|fabq|ipq", colnames (df.bn1), value = TRUE)
phys.var = grep("ext|abds|flex_", colnames (df.bn1), value = TRUE)
fx.var = grep("fitness|endure|balance", colnames (df.bn1), value = TRUE)
others.var = setdiff (names(df.bn1), c(demo.var, pain.var, psych.var, phys.var, fx.var))

pair_var <- expand.grid(from = names (df.bn1),
                        to = names (df.bn1)) %>%
  rownames_to_column()

tiers_keep <- pair_var %>%
  filter (!(grepl (paste0(c(pain.var, psych.var, phys.var, fx.var, demo.var), collapse = "|"),to))) %>%
  filter (!(grepl (paste0(c(pain.var, psych.var, phys.var, fx.var), collapse = "|"),from))) 

  
bl <- anti_join(pair_var, tiers_keep, by = "rowname")  %>%
  filter (from != to) %>%
  select (-rowname) %>%
  mutate_all(as.character) %>% 
  filter (from !="grp") %>% 
  filter (to !="grp") 

latent1 <- expand.grid(from = "pain",
                        to = pain.var)
latent2 <- expand.grid(from = "psych",
                        to = psych.var)
latent3 <- expand.grid(from = "phys",
                        to = phys.var)
latent4 <- expand.grid(from = "funx",
                        to = fx.var)
wl <- latent1 %>%
  bind_rows(latent2) %>%
  bind_rows(latent3)%>%
  bind_rows(latent4)%>%
  mutate_all (as.character)


```

### Structural learning

```{r, eval= FALSE}

####################### Full dataset ####################### ##################

start = bnlearn::bn.fit(bnlearn::empty.graph(names(df.bn)), df.bn)
start$pain = list(coef = 0, sd = 1)
start$phys = list(coef = 0, sd = 1)
start$psych = list(coef = 0, sd = 1)
start$funx = list(coef = 0, sd = 1)

m0 <- bnlearn::structural.em(df.bn, 
                       start = start,
                       maximize = "hc",
                       maximize.args = list(blacklist = bl,
                                            whitelist = wl))


####################### subgroup 1 ####################### ##################
start = bnlearn::bn.fit(bnlearn::empty.graph(names(df.bn1)), df.bn1)
start$pain = list(coef = 0, sd = 1)
start$phys = list(coef = 0, sd = 1)
start$psych = list(coef = 0, sd = 1)
start$funx = list(coef = 0, sd = 1)

m1 <- bnlearn::structural.em(df.bn1, 
                       start = start,
                       maximize = "hc",
                       maximize.args = list(blacklist = bl,
                                            whitelist = wl))



####################### subgroup 2 ####################### ##################
start = bnlearn::bn.fit(bnlearn::empty.graph(names(df.bn2)), df.bn2)
start$pain = list(coef = 0, sd = 1)
start$phys = list(coef = 0, sd = 1)
start$psych = list(coef = 0, sd = 1)
start$funx = list(coef = 0, sd = 1)

m2 <- bnlearn::structural.em(df.bn2, 
                       start = start,
                       maximize = "hc",
                       maximize.args = list(blacklist = bl,
                                            whitelist = wl))

## Plot

par (mfrow = c(2,2), mar = c(1, 1, 1, 1))
plot (m0)
plot (m1)
plot (m2)

```


```{r, eval= FALSE}
doParallel::registerDoParallel(7)
n_boot = 10
############
boot1  =  foreach (B = 1: n_boot) %dopar%{
  set.seed (426)
    boot.sample = df.bn1[sample(nrow(df.bn1)), ]
    
    start = bnlearn::bn.fit(bnlearn::empty.graph(names(boot.sample)), boot.sample)
    start$pain = list(coef = 0, sd = 1)
    start$phys = list(coef = 0, sd = 1)
    start$psych = list(coef = 0, sd = 1)
    start$funx = list(coef = 0, sd = 1)
    bnlearn::structural.em(boot.sample, 
                           start = start,
                           maximize = "hc",
                           maximize.args = list(blacklist = bl,
                                                whitelist = wl))
}
#############\\\\\\\\\\\\\\\\\\\\\\
doParallel::stopImplicitCluster()  


doParallel::registerDoParallel(7)
############
boot2  =  foreach (B = 1: n_boot) %dopar%{
  set.seed (426)
    boot.sample = df.bn2[sample(nrow(df.bn2)), ]
    
    start = bnlearn::bn.fit(bnlearn::empty.graph(names(boot.sample)), boot.sample)
    start$pain = list(coef = 0, sd = 1)
    start$phys = list(coef = 0, sd = 1)
    start$psych = list(coef = 0, sd = 1)
    start$funx = list(coef = 0, sd = 1)
    
    bnlearn::structural.em(boot.sample, 
                           start = start,
                           maximize = "hc",
                           maximize.args = list(blacklist = bl,
                                                whitelist = wl))
}
#############
doParallel::stopImplicitCluster()  


doParallel::registerDoParallel(7)
############
boot3  =  foreach (B = 1: n_boot) %dopar%{
  set.seed (450)
    boot.sample = df.bn3[sample(nrow(df.bn3)), ]
    
    start = bnlearn::bn.fit(bnlearn::empty.graph(names(df.bn2)), df.bn2)
    start$pain = list(coef = 0, sd = 1)
    start$phys = list(coef = 0, sd = 1)
    start$psych = list(coef = 0, sd = 1)
    start$funx = list(coef = 0, sd = 1)
    
    bnlearn::structural.em(boot.sample, 
                           start = start,
                           maximize = "hc",
                           maximize.args = list(blacklist = bl,
                                                whitelist = wl))
}
#############
doParallel::stopImplicitCluster()  

bootstr1 <-  custom.strength(boot1, nodes = names(df.bn1))
avg1 <-  averaged.network(bootstr1, threshold = 0.5)


bootstr2 <-  custom.strength(boot2, nodes = names(df.bn2))
avg2 <-  averaged.network(bootstr2, threshold = 0.7)


bootstr3 <-  custom.strength(boot3, nodes = names(df.bn3))
avg3 <-  averaged.network(bootstr3, threshold = 0.5)

g1 <- strength.plot(avg1, bootstr1)
graph::nodeRenderInfo(g1) <- list(fontsize=30)
Rgraphviz::renderGraph(g1)

g2 <- strength.plot(avg2, bootstr2)
graph::nodeRenderInfo(g2) <- list(fontsize=30)
Rgraphviz::renderGraph(g2)

g3 <- strength.plot(avg3, bootstr3)
graph::nodeRenderInfo(g3) <- list(fontsize=30)
Rgraphviz::renderGraph(g3)

par(mfrow = c(2, 2), mar=c(1,1,1,1))
Rgraphviz::renderGraph(g1)
Rgraphviz::renderGraph(g2)
Rgraphviz::renderGraph(g3)
```
#### Get averaged model

```{r, eval= FALSE}
bootstr <-  custom.strength(boot, nodes = names(df.bn))
avg <-  averaged.network(bootstr, threshold = 0.5)
fit <-  bn.fit (avg, df.bn, method = "mle")
strength.plot(avg, bootstr)
```


## BN to SEM 

### Method 1

```{r}
sem_form0 <- "
  # latent
  physical =~ flex_mob + abds_ms + ext_ms
  pain =~ lbp + legp + duration
  psych =~ ipq + fabq + ases
  funx =~ fitness + endure + balance 
  
  funx ~ age + psych
  psych ~ pain
  physical ~ funx + psych

"

m0 <- lavaan::sem (sem_form0,  data = df_sem, ordered = c("flex_mob", "duration"))
summary (m0)


semPaths (m0, 
          what = "path", 
          layout = "tree3", 
          curvePivot = TRUE, 
          residuals = FALSE,
          intercept = TRUE,
          fixedStyle = c(adjustcolor( "white", alpha.f = 0), 0),
          rotation = 2,
          curve = TRUE,
          freeStyle = c("black", 1))

fitMeasures(m0, fit.measures = c("rmsea", "pvalue", "cfi", "tli", "srmr", "nnfi"))
```

```{r}
sem_form1 <- "
  # latent
  physical =~ flex_mob + abds_ms + ext_ms
  pain =~ lbp + legp + duration
  psych =~ ipq + fabq + ases
  funx =~ fitness + endure + balance 
  
  physical ~ age
  funx ~ physical + psych
  psych ~ physical
  pain ~ funx
"

m1 <- lavaan::sem (sem_form1,  data = df_sem1, ordered = c("flex_mob", "duration"))
summary (m1)

semPaths (m1, 
          what = "path", 
          layout = "tree3", 
          curvePivot = TRUE, 
          residuals = FALSE,
          intercept = TRUE,
          fixedStyle = c(adjustcolor( "white", alpha.f = 0), 0),
          rotation = 2,
          curve = TRUE,
          freeStyle = c("black", 1))

fitMeasures(m1, fit.measures = c("rmsea", "pvalue", "cfi", "tli", "srmr", "nnfi"))
```


```{r}
sem_form2 <- "
  # latent
  physical =~ flex_mob + abds_ms + ext_ms
  pain =~ lbp + legp + duration
  psych =~ ipq + fabq + ases
  funx =~ fitness + endure + balance 
  
  physical ~ age + psych + funx
  psych ~ pain
  funx ~ psych

"

m2 <- lavaan::sem (sem_form2,  data = df_sem2, ordered = c("flex_mob", "duration"))
summary (m2)

semPaths (m2, 
          what = "path", 
          layout = "tree3", 
          curvePivot = TRUE, 
          residuals = FALSE,
          intercept = TRUE,
          fixedStyle = c(adjustcolor( "white", alpha.f = 0), 0),
          rotation = 2,
          curve = TRUE,
          freeStyle = c("black", 1))

fitMeasures(m2, fit.measures = c("rmsea", "pvalue", "cfi", "tli", "srmr", "nnfi"))
```



```{r}
sem_form2 <- "
  # latent
  physical =~ flex_mob + abds_ms + ext_ms
  pain =~ lbp + legp + duration
  psych =~ ipq + fabq + ases
  funx =~ fitness + endure + balance 
  
  physical ~ age + psych + funx
  psych ~ pain
  funx ~ psych

"

m2 <- lavaan::sem (sem_form2,  data = df_sem2, ordered = c("flex_mob", "duration"))
summary (m2)

semPaths (m2, 
          what = "path", 
          layout = "tree3", 
          curvePivot = TRUE, 
          residuals = FALSE,
          intercept = TRUE,
          fixedStyle = c(adjustcolor( "white", alpha.f = 0), 0),
          rotation = 2,
          curve = TRUE,
          freeStyle = c("black", 1))

fitMeasures(m2, fit.measures = c("rmsea", "pvalue", "cfi", "tli", "srmr", "nnfi"))
```




```{r}

m3 <- lavaan::sem (sem_form2,  data = df_sem1, ordered = c("flex_mob", "duration"))
fitMeasures(m3, fit.measures = c("rmsea", "pvalue", "cfi", "tli", "srmr", "nnfi"))
```


# Save data

```{r}
res <- list (m0 = m0,
             m1 = m1,
             m1_2 = m1_2,
             m3 = m3,
             m3_2 = m3_2,
             df = df,
             df3 = df3,
             boot = boot)

saveRDS (res,
         "output/res.RDS")
```

